Constraints on Cosmic-Ray Particle Populations, Magnetic Fields, and Dark Matter from VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS

### 2011-09-01 ms.tex
more typos fixed in the CR/B-analysis part

Moreover, a substantial value of $X_\CR$ can bias hydrostatic cluster masses as
well as the temperature decrement of the CMB in the direction of a galaxy
cluster due to the Sunyaev-Zel'dovich effect which is proportional to the
line-of-sight integral of the thermal electron pressure. This could then
severely jeopardize the use of clusters to determine cosmological
parameters. Comparing X-ray and optical potential profiles in the centers of
galaxy clusters yields an upper limit of 20--30 \% of non-thermal pressure (that
can be composed of CRs, magnetic fields or turbulence) relative to the thermal
gas pressure \cite{2008MNRAS.388.1062C, 2010MNRAS.404.1165C}. An analysis that
compares spatially resolved weak gravitational lensing and hydrostatic X-ray
masses for a sample of 18 galaxy clusters detects a deficit of the hydrostatic
in comparison to the lensing mass of $20 \%$ at $R_{500}$ -- within which the
mean density is 500 times the critical density -- suggesting again a substantial
non-thermal pressure contribution on large scales
\citep{2008MNRAS.384.1567M}. High-energy gamma-ray emission is a complementary
method of constraining the pressure contribution of CRs that is most sensitive
to the cluster core region. However, it assumes that the CR component is fully
mixed with the ICM and almost insensitive in detecting a two-phase structure of
CRs and the thermal ICM.  A fraction $X_\CR$ of only a few percent is required
in order to produce a gamma-ray flux observable with current generation
gamma-ray telescopes rendering this technique at least as sensitive as the
dynamical and hydrostatic method (which are however more general in that they
are sensitive to any non-thermal pressure component).

@@

A third mechanism for gamma-ray production in a galaxy cluster is
self-annihilation of a DM particle, e.g. a weakly interacting massive particle
(WIMP). As already mentioned, about 80\% of the cluster mass is in the form of
dark matter. Recent works \citep{article:EvansFerrerSarkar:2004,
  article:BergstromHooper:2006} have shown that the gamma-ray emissivity from DM
annihilations is expected to be proportional to the square of the DM density
integrated along the line of sight, making clusters of galaxies potentially good
candidates for DM searches \cite{article:PinzkePfrommerBergstrom2009,
  2011ApJ...726L...6C}. This despite their large distances compared to other
common targets for DM searches, such as dwarf spheroidal galaxies
\citep{article:Strigari_etal:2007, article:Acciari_etal:2010,
  article:Aliu_etal:2009} or the Galactic center
\citep{article:Kosack_etal:2004, article:Aharonian_etal:2006,
  article:Aharonian_etal:2009b}.

@@

The Coma cluster of galaxies (ACO 1656) is one of the most thoroughly studied
clusters over all wavelengths \citep{article:Voges_etal:1999}. Located at a
distance of about 100 Mpc \citep[$z=0.023$;][]{article:StrubleRood:1999}, it is
one of the closest massive clusters \citep[$M \sim
10^{15}M_{\odot}$;][]{article:Smith:1983, article:Kubo_etal:2008}. It hosts both
a giant radio halo
\citep{article:Giovannini_etal:1993,article:Thierbach_etal:2003} and peripheral
radio relic, which appears connected to the radio halo with a ``diffuse'' bridge
\citep[see discussion in][]{article:BrownRudnick:2010}. It has been suggested
that the relic is an infall shock \citep{article:Ensslin_etal:1998} and
successively demonstrated by cosmological simulations that model the non-thermal
emission processes \citep{article:PfrommerEnsslinSpringel:2008,
  article:Pfrommer:2008, article:Battaglia_etal:2009,
  article:Skillman_etal:2011}. Extended soft thermal X-ray (SXR) emission is
evident from the ROSAT all-sky survey in the 0.1 to 2.4 keV band
\citep{article:BrielHenryBohringer:1992} and observations with XMM-Newton
\citep{article:Briel_etal:2001} revealed substructure in the X-ray halo with
substantial turbulent pressure support that shows a lower limit of $\sim 10 \%$
of the total intracluster medium pressure \cite{2004A&A...426..387S}. As such it
is a natural candidate for gamma-ray observations.

@@

We note that the overall normalization of the CR and gamma-ray distribution scales with the maximum acceleration efficiency at structure formation shock waves. Following recent observations at supernova remnants \citep{article:Helder_etal:2009} as well as theoretical studies \citep{article:KangJones:2005}, we adopt an optimistic but nevertheless realistic value of this parameter and assume that 50\% of the dissipated energy at strong shocks is injected into CRs while this efficiency rapidly decreases for weaker shocks. Since the vast majority of internal formation shocks (merger and flow shocks) are weak shocks with Mach numbers $M\lesssim3$ \citep[e.g.,][]{article:Ryu_etal:2003}, they do not significantly contribute to the CR population in clusters. Instead, strong shocks during the formation epoch of clusters and strong accretion shocks at the present time (at the boundary of voids and filaments/super-cluster regions) dominate the injection of CRs which are adiabatically transported through the cluster. Hence, these model predictions provide a plausible upper limit for the CR contribution from structure formation shocks in galaxy clusters which can be scaled with the effective acceleration efficiency. Other possible CR sources, such as AGN and starburst-driven galactic winds have been neglected for simplicity which could in principle increase the expected gamma-ray yield.

These cosmological simulations only considered advective transport of CRs by turbulent gas motions which produces centrally enhanced profiles.  However, other means of CR transport such as diffusion and streaming may flatten the CR radial profiles. It has been suggested that advection is dominant only during mergers \citep{article:EnsslinPfrommerMiniatiSubramanian:2011}, in which case relaxed clusters could have a reduced gamma-ray luminosity by up to a factor of five compared with merging clusters. The goal of this and other future work of seeking for gamma-ray emission from clusters is to constrain the acceleration physics and transport properties of CRs.

We adopt the electron density profile for the Coma cluster that has been inferred from ROSAT X-ray observations \citep{article:BrielHenryBohringer:1992} and model the spatial and spectral distribution of CRs according to the model by \citet{article:PinzkePfrommer:2010} that neglects the contribution of cluster galaxies. Figure \ref{fig:spectrum} shows the expected spectrum of Coma, normalized to the total flux within the virial radius. In our further analysis, we use the gamma-ray flux limits of {\em Fermi} in the energy interval of 1-3 GeV which are most constraining.

Figure \ref{fig:xcr} shows the relative CR pressure, $X_{\CR} = \expval{P_{\CR}}/\expval{P_\rmn{th}}$ as a function of radial distance from the Coma cluster center (dashed lines). To compute the CR pressure, we assume a low-energy cutoff of the CR distribution at $T = 0.8\,m_\rmn{p} c^2$ reflecting the high cooling rates at low CR energies. The relative CR pressure rises toward the outer regions on account of the higher efficiency of CR acceleration at the peripheral strong accretion shocks compared to weak central flow shocks. Adiabatic compression of a mixture of CRs and thermal gas disfavours the CR pressure relative to the thermal pressure on account of the softer equation of state of CRs. The weak increase of $X_{\CR}$ toward the core is due to the comparably fast thermal cooling of gas which diminishes the thermal support relative to that in CRs.

@@

\subsection{Simplified analytical model}
To complement the simulation-inspired analysis and to facilitate comparison with
earlier analytical work, we repeat the data analysis with a simplified
analytical model that assumes a power-law CR spectrum and a constant relative CR
pressure, i.e., we adopt the {\em isobaric model of CRs}. Since a priori, the CR
spectral index is unconstrained, we vary it within plausible ranges of
$2.1<\alpha_{\rmn{p}}<2.5$. This spectral range is compatible with the radio
spectral index of the giant radio halo of Coma after accounting for the spectral
steepening at frequencies $\nu\sim5~\rmn{GHz}$ due to the Sunyaev-Zel'dovich
effect \citep{article:PfrommerEnsslin:2004b}\footnote{Assuming a magnetic field
  of 1 $\mu$G, the CR protons responsible for the GHz radio emitting electrons
  have an energy of $\sim100$ GeV and are $\sim$ 20 times less energetic than
  those CR protons responsible for 200-GeV gamma-ray emission.}. Table
\ref{table:constraints} shows the resulting constraints on $X_\CR$. The VERITAS
flux upper limits strongly depend on $\alpha$ due to the comparably large lever
arm from GeV energies (that dominate the CR pressure) to TeV-energies. The
limits within 0.2$^{\circ}$ are most constraining yielding limits on $X_\CR$
between 0.048 and 0.43 (for $\alpha$ varying between 2.1 and 2.5). The {\em
  Fermi} limits depend only weakly on $\alpha$ and are most constraining for an
aperture of 0.4$^{\circ}$.  The best limit on $X_\CR$ of 0.017 is achieved for
$\alpha=2.3$.

@@

\subsection{Minimum gamma-ray flux}
For clusters that host radio halos we are able to derive a minimum gamma-ray flux in the hadronic model of CR interactions. The idea is based on the fact that a stationary distribution of CR electrons loses all its energy to synchrotron radiation for strong magnetic fields ($B \gg B_{\rmn{CMB}} \simeq 3.2 \mu$G) so that the ratio of gamma-ray to synchrotron flux becomes independent of the spatial distribution of CRs and thermal gas \citep{article:Voelk:1989, article:Pohl:1994, article:Pfrommer:2008}, in particular with $\alpha_{\nu}\simeq 1$ as the observed synchrotron spectral index.  Hence we can derive a minimum gamma-ray flux in the hadronic
model
\begin{equation}
\label{eq:Fmin}
\mathcal{F}_{\gamma,\rmn{min}} = \frac{\dps A_{\gamma}}{\dps A_{\nu}}\frac{\dps L_{\nu}}{\dps 4\pi D_{\rmn{lum}}^{2}},
\end{equation}
where $L_{\nu}$ is the observed luminosity of the radio mini-halo,
$D_{\rmn{lum}}$ denotes the luminosity distance to the respective cluster,
$A_\gamma$ and $A_\nu$ are dimensional constants that depend on the hadronic
physics of the interaction \citep{2008MNRAS.385.1211P,
  2008MNRAS.385.1242P}. Lowering the magnetic field would require an increase in
the energy density of CR electrons to reproduce the observed synchrotron
luminosity and thus increase the associated gamma-ray flux.

@@

The results for the minimum gamma-ray flux $F_{\rmn{min}}(>220~\rmn{GeV})$ and the minimum relative CR pressure, $X_{\CR,\,\rmn{min}} = F_{\rmn{min}}/F$ are shown in Table \ref{table:constraints}. Even in the most constraining cases, these are a factor of about 60 below the VERITAS upper limits (adopting $\alpha=2.1$, $<0.2^{\circ}$) and a factor of about 20 below the {\em Fermi} upper limits (adopting $\alpha=2.3$, $<0.4^{\circ}$).

\section{Constraining the Magnetic Field}
\label{sec:B}
We have seen in the previous section that we can obtain an absolute lower limit on the gamma-ray emission in the hadronic model by assuming high magnetic fields, $B>B_\rmn{CMB}$. Hence, we can turn this around by using our upper limit on the gamma-ray emission (and by extension on the CR pressure) to infer a lower limit on the magnetic field that is needed to explain the observed radio emission which, however, assuming the validity of the hadronic model \cite{article:PfrommerEnsslin:2004a}. Lowering the gamma-ray constraint will tighten (increase) the magnetic field limit. Once this gets in conflict with magnetic field measurements by means of other methods, e.g., Faraday rotation measure (RM)\footnote{Generally, Faraday RM analysis of the magnetic-field strength by, e.g., background sources through clusters are degenerate with the magnetic coherence scale and may be biased by the unknown correlation between magnetic and density fluctuations.}, this would challenge the hadronic model of radio halos. The method that we use to constrain the magnetic field inherits a dependence on the assumed spatial structure that we parametrize by
\begin{equation}
\label{eq:B}
B(r) = B_{0} \,\left(\frac{n_{\rmn{e}(r)}}{n_{\rmn{e}(0)}}\right)^{\alpha_B},
\end{equation}
as suggested by Faraday RM studies and numerical magneto-hydrodynamical (MHD) simulations \citep[][and references therein]{article:Bonafede_etal:2010, article:Bonafede_etal:2011}. In fact, the magnetic field in the Coma cluster is among the best constrained, because its proximity permits RM observations of
seven radio-sources located at projected distances of 50 to 1500 kpc from the cluster center. The best-fit model obtains $B_{0} = 4.7^{+0.7}_{-0.8}\,\mu$G and $\alpha_{B} = 0.5^{+0.2}_{-0.1}$ \citep{article:Bonafede_etal:2010}. We aim for constraining the central field strength, $B_{0}$, which depends on the magnetic decline, $\alpha_{B}$, that we vary (within reasonable ranges of $\Delta\alpha_{B}=0.2$ as suggested from the Faraday RM studies) to address this dependence. We proceed as follows:

@@

Table~\ref{table:Bmin} shows the results on the lower limit of the central
magnetic field that range for the VERITAS limits from $B_{0} = 0.4$ to
$1.2\,\mu$G and for the {\em Fermi} limits from $B_{0} = 1$ to $7\,\mu$G. Since
these lower limits on $B_0$ are for most of the parameter space spanned by
$\alpha_B$ and $\alpha$ below the values favored by Faraday RM, the hadronic
model is still viable. The only exception is the parameter combination of
$alpha_B=0.7$ and $\alpha=2.5$ which starts to be in (weak) tension with Faraday
RM data. This indicates the power of future gamma-ray observations of Coma in
constraining parameters of the hadronic model.

In addition there are a few interesting remarks in order. (1) For the VERITAS
limits, the hardest CR spectral indices are most constraining because of the
tightest upper limits. The opposite is true for the {\em Fermi} upper limits
which probe the region around the pion decay bump characterized by a shape
parameter that depends inversely on the CR spectral index
\cite{article:PfrommerEnsslin:2004b}.  (2) For a steeper magnetic decline
(larger $\alpha_{B}$), the CR number density needs to be larger to match the
observed radio emission profiles which would yield a higher gamma-ray flux so
that the upper limits are more constraining. This implies tighter lower limits
for $B_{0}$. However, the largest choice of $\alpha_B=0.7$, which represents a
flux-freezing magnetic field, is possibly too steep in the outer parts as it
neglects the contributions from turbulent dynamos in these dynamically active
regions with vorticity injection at corrugated structure formation shocks
\cite{article:PfrommerJones:2011}. (3) Interestingly, in all cases, we use the
0.4$^{\circ}$ limit that is always most constraining for our choice of magnetic
field configurations.  This is because we need a substantially increasing
relative CR pressure profile to match the observed radio profiles (for a given
magnetic realization); hence that CR realization produces a larger flux within
0.4$^{\circ}$ in comparison to the ``isobaric CR model'' assumption ($X_{\CR} =
\rmn{const.}$) for which the 0.2$^{\circ}$ upper limits are more constraining;
hence the constraints on $B_{0}$ get better for the larger integration
area. Physically, this may be achieved by considering CR transport processes
such as streaming which produces an increasing $X_{\CR}$-profile due to the
larger available phase space in the outer cluster regions.

@@

Within this pragmatic approach we employ two different directions. First,
adopting a high magnetic field everywhere in the cluster ($B\gg B_\rmn{CMB}$)
yields the minimum gamma-ray flux in the hadronic model of radio halos which we
find to be a factor of 20 (60) below the most constraining flux upper limits of
{\em Fermi} (VERITAS). Second, by matching the radio emission profile
(i.e. fixing the radial CR profile for a given magnetic field model) and by
requiring the pion-decay gamma-ray flux to match the flux upper limits
(i.e. fixing the normalization of the CR distribution) we obtain lower limits on
the magnetic field distribution under consideration. Our limits of the central
magnetic field range from $B_{0} = 0.4$ to $1.2\,\mu$G (for VERITAS limits) and
from $B_{0} = 1$ to $7\,\mu$G (for {\em Fermi} limits). Since these lower limits
on $B_0$ are below the values favored by Faraday RM, the hadronic model is still
viable.

@@

\begin{figure*}
\begin{center}
\scalebox{1.0}{\plotone{f5.eps}}
\end{center}
\caption{CR-to-thermal pressure, $X_{\mathrm{CR}}=\expval{P_{\CR}}/ \expval{P_{\mathrm{th}}}$, as a function of radial distance from the center of the Coma cluster. Model predictions are shown by dashed lines \citep{article:PinzkePfrommer:2010} while these are scaled to match the most constraining VERITAS upper limits within 0.2$^\circ$ (solid). We compare differential $X_\CR$-profiles (red) to integrated profiles $X_{\CR}(<R/R_{\rmn{vir}})$ (blue) which we use to compare to the upper limits.}
\label{fig:xcr}
\end{figure*}

>>>>

Moreover, a substantial value of $X_\CR$ can bias hydrostatic cluster masses as
well as the temperature decrement of the CMB in the direction of a galaxy
cluster due to the Sunyaev-Zel'dovich effect which is proportional to the
line-of-sight integral of the thermal electron pressure. This could then
severely jeopardize the use of clusters to determine cosmological
parameters. Comparing X-ray and optical potential profiles in the centers of
galaxy clusters yields an upper limit of 20--30 \% of non-thermal pressure (that
can be composed of CRs, magnetic fields or turbulence) relative to the thermal
gas pressure \citep{2008MNRAS.388.1062C, 2010MNRAS.404.1165C}. An analysis that
compares spatially resolved weak gravitational lensing and hydrostatic X-ray
masses for a sample of 18 galaxy clusters detects a deficit of the hydrostatic
in comparison to the lensing mass of $20 \%$ at $R_{500}$ -- within which the
mean density is 500 times the critical density -- suggesting again a substantial
non-thermal pressure contribution on large scales
\citep{2008MNRAS.384.1567M}. High-energy gamma-ray emission is a complementary
method of constraining the pressure contribution of CRs that is most sensitive
to the cluster core region. However, it assumes that the CR component is fully
mixed with the ICM and almost insensitive in detecting a two-phase structure of
CRs and the thermal ICM.  A fraction $X_\CR$ of only a few percent is required
in order to produce a gamma-ray flux observable with current generation
gamma-ray telescopes rendering this technique at least as sensitive as the
dynamical and hydrostatic method (which are however more general in that they
are sensitive to any non-thermal pressure component).

@@

A third mechanism for gamma-ray production in a galaxy cluster is
self-annihilation of a DM particle, e.g. a weakly interacting massive particle
(WIMP). As already mentioned, about 80\% of the cluster mass is in the form of
dark matter. Recent works \citep{article:EvansFerrerSarkar:2004,
  article:BergstromHooper:2006} have shown that the gamma-ray emissivity from DM
annihilations is expected to be proportional to the square of the DM density
integrated along the line of sight, making clusters of galaxies potentially good
candidates for DM searches \citep{article:PinzkePfrommerBergstrom2009,
  2011ApJ...726L...6C}. This despite their large distances compared to other
common targets for DM searches, such as dwarf spheroidal galaxies
\citep{article:Strigari_etal:2007, article:Acciari_etal:2010,
  article:Aliu_etal:2009} or the Galactic center
\citep{article:Kosack_etal:2004, article:Aharonian_etal:2006,
  article:Aharonian_etal:2009b}.

@@

The Coma cluster of galaxies (ACO 1656) is one of the most thoroughly studied
clusters over all wavelengths \citep{article:Voges_etal:1999}. Located at a
distance of about 100 Mpc \citep[$z=0.023$;][]{article:StrubleRood:1999}, it is
one of the closest massive clusters \citep[$M \sim
10^{15}M_{\odot}$;][]{article:Smith:1983, article:Kubo_etal:2008}. It hosts both
a giant radio halo
\citep{article:Giovannini_etal:1993,article:Thierbach_etal:2003} and peripheral
radio relic, which appears connected to the radio halo with a ``diffuse'' bridge
\citep[see discussion in][]{article:BrownRudnick:2010}. It has been suggested
that the relic is an infall shock \citep{article:Ensslin_etal:1998} and
successively demonstrated by cosmological simulations that model the non-thermal
emission processes \citep{article:PfrommerEnsslinSpringel:2008,
  article:Pfrommer:2008, article:Battaglia_etal:2009,
  article:Skillman_etal:2011}. Extended soft thermal X-ray (SXR) emission is
evident from the ROSAT all-sky survey in the 0.1 to 2.4 keV band
\citep{article:BrielHenryBohringer:1992} and observations with XMM-Newton
\citep{article:Briel_etal:2001} revealed substructure in the X-ray halo with
substantial turbulent pressure support that shows a lower limit of $\sim 10 \%$
of the total intracluster medium pressure \citep{2004A&A...426..387S}. As such it
is a natural candidate for gamma-ray observations.

@@

The overall normalization of the CR and gamma-ray distribution scales with the maximum acceleration efficiency at structure formation shock waves. Following recent observations at supernova remnants \citep{article:Helder_etal:2009} as well as theoretical studies \citep{article:KangJones:2005}, we adopt an optimistic but nevertheless realistic value of this parameter and assume that 50\% of the dissipated energy at strong shocks is injected into CRs while this efficiency rapidly decreases for weaker shocks. Since the vast majority of internal formation shocks (merger and flow shocks) are weak shocks with Mach numbers $M\lesssim3$ \citep[e.g.,][]{article:Ryu_etal:2003}, they do not significantly contribute to the CR population in clusters. Instead, strong shocks during the formation epoch of clusters and strong accretion shocks at the present time (at the boundary of voids and filaments/super-cluster regions) dominate the injection of CRs which are adiabatically transported through the cluster. Hence, these model predictions provide a plausible upper limit for the CR contribution from structure formation shocks in galaxy clusters which can be scaled with the effective acceleration efficiency. Other possible CR sources, such as AGN and starburst-driven galactic winds have been neglected for simplicity which could in principle increase the expected gamma-ray yield.

These cosmological simulations only considered advective transport of CRs by turbulent gas motions which produces centrally enhanced profiles.  However, other means of CR transport such as diffusion and streaming may flatten the CR radial profiles. It has been suggested that advection is dominant only during mergers, in which case relaxed clusters could have a reduced gamma-ray luminosity by up to a factor of five compared with merging clusters \citep{article:EnsslinPfrommerMiniatiSubramanian:2011}. The goal of this and other future work of seeking for gamma-ray emission from clusters is to constrain the acceleration physics and transport properties of CRs.

We adopt the electron density profile for the Coma cluster that has been inferred from ROSAT X-ray observations \citep{article:BrielHenryBohringer:1992} and model the spatial and spectral distribution of CRs according to the model by \citet{article:PinzkePfrommer:2010} that neglects the contribution of cluster galaxies. Figure \ref{fig:spectrum} shows the expected spectrum of Coma, normalized to the total flux within the virial radius. In our further analysis, we use the gamma-ray flux limits of {\em Fermi} in the energy interval of 1--3 GeV which are most constraining as well as the gamma-ray flux limits of VERITAS in the energy range above 220 GeV.

Figure \ref{fig:xcr} shows the relative CR pressure, $X_{\CR} = \expval{P_{\CR}}/\expval{P_\rmn{th}}$ as a function of radial distance, $R$, from the Coma cluster center (dashed red lines) and contained within $R$ (dashed blue lines). To compute the CR pressure, we assume a low-momentum cutoff of the CR distribution at $q = 0.8\,m_p c$ as found in the cosmological simulations which reflects the high cooling rates at low CR energies. The relative CR pressure rises toward the outer regions on account of the higher efficiency of CR acceleration at the peripheral strong accretion shocks compared to weak central flow shocks. Adiabatic compression of a mixture of CRs and thermal gas disfavours the CR pressure relative to the thermal pressure on account of the softer equation of state of CRs. The weak increase of $X_{\CR}$ toward the core is due to the comparably fast thermal cooling of gas which diminishes the thermal support relative to that in CRs.

@@

\subsection{Simplified analytical model}
To complement the simulation-inspired analysis and to facilitate comparison with earlier analytical work, we repeat the data analysis with a simplified analytical model that assumes a power-law CR spectrum and a constant relative CR pressure, i.e., we adopt the {\em isobaric model of CRs}. To be independent of additional assumptions and in line with earlier work in the literature, we do not impose a low-momentum cutoff on the CR distribution function, i.e., we adopt $q=0$. Since a priori, the CR spectral index is unconstrained, we vary it within plausible ranges of $2.1<\alpha_{\rmn{p}}<2.5$. This spectral range is compatible with the radio spectral index of the giant radio halo of Coma after accounting for the spectral steepening at frequencies $\nu\sim5~\rmn{GHz}$ due to the Sunyaev-Zel'dovich effect \citep{article:PfrommerEnsslin:2004b}\footnote{Assuming a magnetic field of 1 $\mu$G, the CR protons responsible for the GHz radio emitting electrons have an energy of $\sim100$ GeV and
  are $\sim$ 20 times less energetic than those CR protons responsible for 200-GeV gamma-ray emission.}. Table \ref{table:constraints} shows the resulting constraints on $X_\CR$. The VERITAS flux upper limits strongly depend on $\alpha$ due to the comparably large lever arm from GeV energies (that dominate the CR pressure) to TeV-energies. The limits within 0.2$^{\circ}$ are most constraining yielding limits on $X_\CR$ between 0.048 and 0.43 (for $\alpha$ varying between 2.1 and 2.5). The {\em Fermi} limits depend only weakly on $\alpha$ and are most constraining for an aperture of 0.4$^{\circ}$.  The best limit on $X_\CR$ of 0.017 is achieved for $\alpha=2.3$.

@@

\subsection{Minimum gamma-ray flux}
For clusters that host radio halos we are able to derive a minimum gamma-ray flux in the hadronic model of CR interactions. The idea is based on the fact that a stationary distribution of CR electrons loses all its energy to synchrotron radiation for strong magnetic fields ($B \gg B_{\rmn{CMB}} \simeq 3.2 \mu$G) so that the ratio of gamma-ray to synchrotron flux becomes independent of the spatial distribution of CRs and thermal gas \citep{article:Voelk:1989, article:Pohl:1994, article:Pfrommer:2008}, in particular with $\alpha_{\nu}\simeq 1$ as the observed synchrotron spectral index.  Hence we can derive a minimum gamma-ray flux in the hadronic model
\begin{equation}
\label{eq:Fmin}
\mathcal{F}_{\gamma,\rmn{min}} = \frac{\dps A_{\gamma}}{\dps A_{\nu}}\frac{\dps L_{\nu}}{\dps 4\pi D_{\rmn{lum}}^{2}},
\end{equation}
where $L_{\nu}$ is the observed luminosity of the radio mini-halo, $D_{\rmn{lum}}$ denotes the luminosity distance to the respective cluster, $A_\gamma$ and $A_\nu$ are dimensional constants that depend on the hadronic physics of the interaction \citep{article:Pfrommer:2008, Pfrommer_etal:2008}. Lowering the magnetic field would require an increase in the energy density of CR electrons to reproduce the observed synchrotron luminosity and thus increase the associated gamma-ray flux.

@@

The results for the minimum gamma-ray flux $F_{\rmn{min}}(>220~\rmn{GeV})$ and the minimum relative CR pressure, $X_{\CR,\,\rmn{min}} = F_{\rmn{min}}/\tilde{F}$ are shown in Table \ref{table:constraints}, where $F_\rmn{iso} = X_\CR\,\tilde{F}$ is the gamma-ray flux in the {\em isobaric model of CRs} introduced in the previous section. Even in the most constraining cases, these are a factor of about 60 below the VERITAS upper limits (adopting $\alpha=2.1$, $<0.2^{\circ}$) and a factor of about 20 below the {\em Fermi} upper limits (adopting $\alpha=2.3$, $<0.4^{\circ}$).

\section{Constraining the Magnetic Field}
\label{sec:B}
We have seen in the previous section that we can obtain an absolute lower limit on the gamma-ray emission in the hadronic model by assuming high magnetic fields, $B>B_\rmn{CMB}$. Hence, we can turn this around by using our upper limit on the gamma-ray emission (and by extension on the CR pressure) to infer a lower limit on the magnetic field that is needed to explain the observed radio emission which, however, assuming the validity of the hadronic model \citep{article:PfrommerEnsslin:2004a}. Lowering the gamma-ray constraint will tighten (increase) the magnetic field limit. Once this gets in conflict with magnetic field measurements by means of other methods, e.g., Faraday rotation measure (RM)\footnote{Generally, Faraday RM analysis of the magnetic-field strength by, e.g., background sources through clusters are degenerate with the magnetic coherence scale and may be biased by the unknown correlation between magnetic and density fluctuations.}, this would challenge the hadronic model of radio halos. The method that we use to constrain the magnetic field inherits a dependence on the assumed spatial structure that we parametrize by
\begin{equation}
\label{eq:B}
B(r) = B_{0} \,\left(\frac{n_{\rmn{e}(r)}}{n_{\rmn{e}(0)}}\right)^{\alpha_B},
\end{equation}
as suggested by Faraday RM studies and numerical magneto-hydrodynamical (MHD) simulations \citep[][and references therein]{article:Bonafede_etal:2010, article:Bonafede_etal:2011}. In fact, the magnetic field in the Coma cluster is among the best constrained, because its proximity permits RM observations of
seven radio-sources located at projected distances of 50 to 1500 kpc from the cluster center. The best-fit model obtains $B_{0} = 4.7^{+0.7}_{-0.8}\,\mu$G and $\alpha_{B} = 0.5^{+0.2}_{-0.1}$ \citep{article:Bonafede_etal:2010}. We aim for constraining the central field strength, $B_{0}$, which depends on the magnetic decline, $\alpha_{B}$, that we vary (within reasonable ranges of $\Delta\alpha_{B}=0.2$ as suggested from the Faraday RM studies) to address this dependence. We proceed as follows:

@@

Table~\ref{table:Bmin} shows the results on the lower limit of the central magnetic field that range for the VERITAS limits from $B_{0} = 0.4$ to $1.2\,\mu$G and for the {\em Fermi} limits from $B_{0} = 1$ to $7\,\mu$G. Since these lower limits on $B_0$ are below the values favored by Faraday RM for most of the parameter space spanned by $\alpha_B$ and $\alpha$, the hadronic model is still viable. The only exception is the parameter combination of $\alpha_B=0.7$ and $\alpha=2.5$ which starts to be in (weak) tension with Faraday RM data. This indicates the power of future gamma-ray observations of Coma in constraining parameters of the hadronic model.

In addition there are a few interesting remarks in order. (1) For the VERITAS limits, the hardest CR spectral indices are most constraining because of the tightest upper limits. The opposite is true for the {\em Fermi} upper limits which probe the region around the pion decay bump characterized by a shape parameter that depends inversely on the CR spectral index \citep{article:PfrommerEnsslin:2004b}.  (2) For a steeper magnetic decline (larger $\alpha_{B}$), the CR number density needs to be larger to match the observed radio emission profiles which would yield a higher gamma-ray flux so that the upper limits are more constraining. This implies tighter lower limits for $B_{0}$. However, the largest choice of $\alpha_B=0.7$, which represents a flux-freezing magnetic field, is possibly too steep in the outer parts as it neglects the contributions from turbulent dynamos in these dynamically active regions with vorticity injection at corrugated structure formation shocks
\citep{article:PfrommerJones:2011}. (3) Interestingly, in all cases, we use the 0.4$^{\circ}$ limit that is always most constraining for our choice of magnetic field configurations.  This is because we need a substantially increasing relative CR pressure profile to match the observed radio profiles (for a given magnetic realization); hence that CR realization produces a larger flux within 0.4$^{\circ}$ in comparison to the ``isobaric CR model'' assumption ($X_{\CR} = \rmn{const.}$) for which the 0.2$^{\circ}$ upper limits are more constraining; hence the constraints on $B_{0}$ get better for the larger integration area. Physically, this may be achieved by considering CR transport processes such as streaming which produces an increasing $X_{\CR}$-profile due to the larger available phase space in the outer cluster regions.

@@

Within this pragmatic approach we employ two different directions. First, adopting a high magnetic field everywhere in the cluster ($B\gg B_\rmn{CMB}$) yields the minimum gamma-ray flux in the hadronic model of radio halos which we find to be a factor of 20 (60) below the most constraining flux upper limits of {\em Fermi} (VERITAS). Second, by matching the radio emission profile (i.e., fixing the radial CR profile for a given magnetic field model) and by requiring the pion-decay gamma-ray flux to match the flux upper limits (i.e., fixing the normalization of the CR distribution) we obtain lower limits on the magnetic field distribution under consideration. Our limits of the central magnetic field range from $B_{0} = 0.4$ to $1.2\,\mu$G (for VERITAS gamma-ray flux limits) and from $B_{0} = 1$ to $7\,\mu$G (for {\em Fermi} gamma-ray flux limits). Since these lower limits on $B_0$ are below the values favored by Faraday RM, the hadronic model is still viable.

@@

\begin{figure*}
\begin{center}
\scalebox{1.0}{\plotone{f5.eps}}
\end{center}
\caption{CR-to-thermal pressure, $X_{\mathrm{CR}}=\expval{P_{\CR}}/ \expval{P_{\mathrm{th}}}$, as a function of radial distance from the center of the Coma cluster. Model predictions are shown by dashed lines \citep{article:PinzkePfrommer:2010} while these are scaled to match the most constraining VERITAS upper limits within 0.2$^\circ$ (solid). We compare differential $X_\CR$-profiles (red) to integrated profiles $X_{\CR}(<R/R_{\rmn{vir}})=\int_0^R P_\CR\, dV / \int_0^R P_{\rmn{th}}\, dV$ (blue) which we use to compare to the upper limits.}
\label{fig:xcr}
\end{figure*}
